```{r modelsp, include=FALSE, eval=T}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(cmdstanr)
library(rstan)
library(bayesplot)
library(loo)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F, eval=T)
```

# Species integration

In this chapter, I looked for the best model to integrate species information based on quality of fit, cross-validation (LOO), and prediction (RMSEP). The result is summarised in Table \@ref(tab:spchoicetab).

## Data

I used reduced data to explore the model form. 
I focused on trees at 20 meters from any plot edges for neighbourhood effect.
I used only recruited trees in the censuses with at least 10 measurements of diameter at breast height (DBH, cm).
I used only species with at least 10 trees following previous requirements.
And, I randomly selected 10 trees among 9 species for lightweight training data.
I further selected a random diameter measure for each individual as an evaluation data not used for model fitting (Fig. \@ref(fig:mmmdataspfig)).

```{r datasp, eval=F}
guyafor <- DBI::dbConnect(RSQLite::SQLite(), dbname = "data/guyafor.sql")
trees <- tbl(guyafor, "inventory") %>% 
  filter(Forest == "Paracou") %>% 
  collect() %>% 
  filter(Xfield > 20, Xfield < 250-20, Yfield > 20, Yfield < 250-20) %>% 
  mutate(species = paste(Genre, Espece)) %>% 
  filter(!grepl("Indet", species)) %>% 
  mutate(DBH = CircCorr/pi)
DBI::dbDisconnect(guyafor) ; rm(guyafor)
```

```{r mdatasp, eval=F}
trees %>% 
  group_by(Plot) %>% 
  mutate(StartYear = min(CensusYear)) %>% 
  group_by(idTree) %>% 
  arrange(CensusYear) %>% 
  filter(first(CensusYear) > StartYear) %>% 
  filter(first(DBH) < 12) %>% 
  filter(last(DBH) > first(DBH)) %>% 
  filter(n() > 10) %>% 
  mutate(Year = CensusYear - first(CensusYear)) %>%
  mutate(DBHprev = lag(DBH)) %>% 
  mutate(Deltat = CensusYear - lag(CensusYear)) %>% 
  mutate(DBHprev = ifelse(is.na(DBHprev), 10, DBHprev)) %>% 
  mutate(Deltat = ifelse(is.na(Deltat), 1, DBHprev)) %>% 
  ungroup() %>% 
  group_by(species) %>% 
  filter(length(unique(idTree)) > 10) %>% 
  ungroup() %>% 
  filter(species %in% sample(unique(species), 9)) %>% 
  group_by(species) %>% 
  filter(idTree %in% sample(unique(idTree), 10)) %>%
  ungroup() %>% 
  mutate(ind = as.numeric(as.factor(as.character(idTree)))) %>% 
  mutate(sp = as.numeric(as.factor(as.character(species)))) %>% 
  vroom::vroom_write("save/species/trees_species.tsv")
vroom::vroom("save/species/trees_species.tsv") %>% 
  group_by(idTree) %>% 
  sample_n(1) %>% 
  vroom::vroom_write("save/species/eval_species.tsv")
vroom::vroom("save/species/trees_species.tsv") %>% 
  anti_join(vroom::vroom("save/species/eval_species.tsv")) %>% 
  vroom::vroom_write("save/species/mdata_species.tsv")
```

```{r mmmdataspfig, fig.cap="Tree diameter trajectories in training data. Color represent individuals and the red point the data used for evaluation."}
ggplot(vroom::vroom("save/species/mdata_species.tsv"),
       aes(Year, DBH, col = as.factor(ind))) +
  geom_point() +
  geom_smooth(se = F) +
  geom_point(data = vroom::vroom("save/species/eval_species.tsv"), col = "red", size = 2) +
  scale_color_discrete(guide = "none") +
  facet_wrap(~ species)
```

## Individual fixed

We used the selected model, *i.e.* Gompertz model [@Herault2011], with all parameters as individual fixed effects:

$$  DBH_{t,i} \sim \mathcal N  (10 + gmax_i \times \sum _{y=1|DBH_{t=0}} ^{y=t} exp(-\frac12.[\frac{log(\frac{DBH_{t,i}}{100.Dopt_i})}{Ks_i}]^2)), \sigma)$$

The model correctly converged ($\hat R < 1.05$) with an acceptable but marked correlation between $dopt$ and $ks$.
$gmax$ posteriors have logical uncertainty but are varying widely among individuals.
Finally predictions of the diameter trajectories by the model are good and realistic.
Nevertheless, some posteriors seems to show bimodality.

```{r indfixfit, eval=F}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
indfix <- cmdstan_model("model/indfix.stan")
fit <- indfix$sample(
  data = list(
    N = nrow(mdata),
    I = max(mdata$ind),
    S = max(mdata$sp),
    Y = max(max(mdata$Year), max(eval$Year))+1,
    year = mdata$Year+1,
    ind = mdata$ind,
    sp = mdata$sp,
    dbh = mdata$DBH,
    Np = nrow(eval),
    dbhp = eval$DBH,
    indp = eval$ind,
    spp = eval$sp,
    yearp = eval$Year+1),
  chains = 2, 
  parallel_chains = 4,
  refresh = 200,
  save_warmup = F
)
fit_indfix <- rstan::read_stan_csv(fit$output_files())
save(fit_indfix, file = "save/species/indfix.Rdata") 
```

```{r indfixmsampling}
load("save/species/indfix.Rdata")  
```

```{r indfixrhat}
mcmc_rhat(rhat(fit_indfix)) 
```

```{r indfixtrace}
mcmc_trace(as.array(fit_indfix, pars = c("gmax[1]", "dopt[1]", "ks[1]", "sigma")))
```

```{r indfixpairs}
mcmc_pairs(as.array(fit_indfix, pars = c("gmax[1]", "dopt[1]", "ks[1]", "sigma"))) 
```

```{r indfixalpha}
mcmc_intervals(as.array(fit_indfix, pars = "gmax"))
```

```{r indfixpreds}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
dbhtoday <- function(t, gmax, dopt, ks){
  dbh <- 10
  for(i in 1:t)
    dbh <- dbh + gmax*exp(-0.5* (log(dbh / (100*dopt)) / ks)^2)
  return(dbh)  
}
pars <- as.data.frame(fit_indfix, pars = c('gmax', "dopt", "ks")) %>% 
  reshape2::melt(variable.name = "parameter") %>% 
  group_by(parameter) %>% 
  summarise(value = median(value)) %>% 
  separate(parameter, c("parameter", "ind"), convert = T) %>% 
  reshape2::dcast(ind ~ parameter)
preds <- pars %>% 
  mutate(Year = list(0:max(eval$Year))) %>% 
  unnest(Year) %>% 
  rowwise() %>% 
  mutate(DBH = dbhtoday(Year+1, gmax, dopt, ks))
ggplot(left_join(mdata, pars), aes(Year, DBH, col = log(gmax), group = ind)) +
  geom_point() +
  geom_point(data = eval, col = "red", size = 2) +
  geom_line(data = preds) +
  viridis::scale_color_viridis() 
```

## Individual random

We used the selected model, *i.e.* Gompertz model [@Herault2011], with $gmax$ as individual fixed effects and $Dopt$ and $Ks$ as individual random effects centered on species fixed effects:

$$  DBH_{t,i,s} \sim \mathcal N  (10 + gmax_i \times \sum _{y=1|DBH_{t=0}} ^{y=t} exp(-\frac12.[\frac{log(\frac{DBH_{t,i}}{100.Dopt_i})}{Ks_i}]^2)), \sigma) \\| Dopt_i \sim \mathcal N(Dopt_s,\sigma_D), Ks_i \sim \mathcal N(Ks_s,\sigma_K) $$

The model correctly converged ($\hat R < 1.05$) with an acceptable but marked correlation between $dopt$ and $ks$.
$gmax$ posteriors have logical uncertainty but are varying widely among individuals.
Finally predictions of the diameter trajectories by the model are good and realistic.

```{r indrandfit, eval=F}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
indrand <- cmdstan_model("model/indrand.stan")
fit <- indrand$sample(
  data = list(
    N = nrow(mdata),
    I = max(mdata$ind),
    S = max(mdata$sp),
    Y = max(max(mdata$Year), max(eval$Year))+1,
    year = mdata$Year+1,
    ind = mdata$ind,
    sp = mdata$sp,
    indsp = arrange(unique(mdata[c("ind", "sp")]), ind)$sp,
    dbh = mdata$DBH,
    Np = nrow(eval),
    dbhp = eval$DBH,
    indp = eval$ind,
    yearp = eval$Year+1),
  chains = 2, 
  parallel_chains = 4,
  refresh = 200,
  save_warmup = F
)
fit_indrand <- rstan::read_stan_csv(fit$output_files())
save(fit_indrand, file = "save/species/indrand.Rdata") 
```

```{r indrandmsampling}
load("save/species/indrand.Rdata")  
```

```{r indrandrhat}
mcmc_rhat(rhat(fit_indrand)) 
```

```{r indrandtrace}
mcmc_trace(as.array(fit_indrand, pars = c("gmax[1]", "sigma", "dopt_s[1]", "ks_s[1]", "sigmaD", "sigmaK")))
```

```{r indrandpairs}
mcmc_pairs(as.array(fit_indrand, pars = c("gmax[1]", "sigma", "dopt_s[1]", "ks_s[1]", "sigmaD", "sigmaK"))) 
```

```{r indrandalpha}
mcmc_intervals(as.array(fit_indrand, pars = "gmax"))
```

```{r indrandpreds}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
dbhtoday <- function(t, gmax, dopt, ks){
  dbh <- 10
  for(i in 1:t)
    dbh <- dbh + gmax*exp(-0.5* (log(dbh / (100*dopt)) / ks)^2)
  return(dbh)  
}
pars <- as.data.frame(fit_indrand, pars = c('gmax', "dopt", "ks")) %>% 
  reshape2::melt(variable.name = "parameter") %>% 
  group_by(parameter) %>% 
  summarise(value = median(value)) %>% 
  separate(parameter, c("parameter", "ind"), convert = T) %>% 
  reshape2::dcast(ind ~ parameter)
preds <- pars %>% 
  mutate(Year = list(0:max(eval$Year))) %>% 
  unnest(Year) %>% 
  rowwise() %>% 
  mutate(DBH = dbhtoday(Year+1, gmax, dopt, ks))
ggplot(left_join(mdata, pars), aes(Year, DBH, col = log(gmax), group = ind)) +
  geom_point() +
  geom_point(data = eval, col = "red", size = 2) +
  geom_line(data = preds) +
  viridis::scale_color_viridis() 
```

## Species fixed

We used the selected model, *i.e.* Gompertz model [@Herault2011], with $gmax$ as individual fixed effects and $Dopt$ and $Ks$ as species fixed effects:

$$  DBH_{t,i,s} \sim \mathcal N  (10 + gmax_i \times \sum _{y=1|DBH_{t=0}} ^{y=t} exp(-\frac12.[\frac{log(\frac{DBH_{t,i}}{100.Dopt_s})}{Ks_s}]^2)), \sigma)$$

The model correctly converged ($\hat R < 1.05$) with an acceptable but marked correlation between $dopt$ and $ks$.
$gmax$ posteriors have logical uncertainty but are varying widely among individuals.
Finally predictions of the diameter trajectories by the model are good and realistic.

```{r spfixfit, eval=F}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
spfix <- cmdstan_model("model/spfix.stan")
fit <- spfix$sample(
  data = list(
    N = nrow(mdata),
    I = max(mdata$ind),
    S = max(mdata$sp),
    Y = max(max(mdata$Year), max(eval$Year))+1,
    year = mdata$Year+1,
    ind = mdata$ind,
    sp = mdata$sp,
    indsp = arrange(unique(mdata[c("ind", "sp")]), ind)$sp,
    dbh = mdata$DBH,
    Np = nrow(eval),
    dbhp = eval$DBH,
    indp = eval$ind,
    yearp = eval$Year+1),
  chains = 2, 
  parallel_chains = 4,
  refresh = 200,
  save_warmup = F
)
fit_spfix <- rstan::read_stan_csv(fit$output_files())
save(fit_spfix, file = "save/species/spfix.Rdata") 
```

```{r spfixmsampling}
load("save/species/spfix.Rdata")  
```

```{r spfixrhat}
mcmc_rhat(rhat(fit_spfix)) 
```

```{r spfixtrace}
mcmc_trace(as.array(fit_spfix, pars = c("gmax[1]", "sigma", "dopt_s[1]", "ks_s[1]")))
```

```{r spfixpairs}
mcmc_pairs(as.array(fit_spfix, pars = c("gmax[1]", "sigma", "dopt_s[1]", "ks_s[1]"))) 
```

```{r spfixalpha}
mcmc_intervals(as.array(fit_spfix, pars = "gmax"))
```

```{r spfixpreds}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
dbhtoday <- function(t, gmax, dopt, ks){
  dbh <- 10
  for(i in 1:t)
    dbh <- dbh + gmax*exp(-0.5* (log(dbh / (100*dopt)) / ks)^2)
  return(dbh)  
}
parsi <- as.data.frame(fit_spfix, pars = c("gmax")) %>% 
  reshape2::melt(variable.name = "parameter") %>% 
  group_by(parameter) %>% 
  summarise(value = median(value)) %>% 
  separate(parameter, c("parameter", "ind"), convert = T) %>% 
  reshape2::dcast(ind ~ parameter)
parss <- as.data.frame(fit_spfix, pars = c("dopt_s", "ks_s")) %>% 
  reshape2::melt(variable.name = "parameter") %>% 
  group_by(parameter) %>% 
  summarise(value = median(value)) %>% 
  separate(parameter, c("parameter", "X1",  "sp"), convert = T) %>% 
  reshape2::dcast(sp ~ parameter)
preds <- select(mdata, ind, sp) %>% 
  unique() %>% 
  left_join(parsi) %>% 
  left_join(parss) %>% 
  mutate(Year = list(0:max(eval$Year))) %>% 
  unnest(Year) %>% 
  rowwise() %>% 
  mutate(DBH = dbhtoday(Year+1, gmax, dopt, ks))
ggplot(left_join(mdata, pars), aes(Year, DBH, col = log(gmax), group = ind)) +
  geom_point() +
  geom_point(data = eval, col = "red", size = 2) +
  geom_line(data = preds) +
  viridis::scale_color_viridis() 
```

## Species random

We used the selected model, *i.e.* Gompertz model [@Herault2011], with $gmax$ as individual fixed effects and $Dopt$ and $Ks$ as species random effects centered on community fixed effects:

$$  DBH_{t,i,s} \sim \mathcal N  (10 + gmax_i \times \sum _{y=1|DBH_{t=0}} ^{y=t} exp(-\frac12.[\frac{log(\frac{DBH_{t,i}}{100.Dopt_s})}{Ks_s}]^2)), \sigma) \\| Dopt_i \sim \mathcal N(Dopt,\sigma_K), Ks \sim \mathcal N(Ks,\sigma_D) $$

The model correctly converged ($\hat R < 1.05$) with an acceptable but marked correlation between $dopt$ and $ks$.
$gmax$ posteriors have logical uncertainty but are varying widely among individuals.
Finally predictions of the diameter trajectories by the model are good and realistic.

```{r sprandfit, eval=F}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
sprand <- cmdstan_model("model/sprand.stan")
fit <- sprand$sample(
  data = list(
    N = nrow(mdata),
    I = max(mdata$ind),
    S = max(mdata$sp),
    Y = max(max(mdata$Year), max(eval$Year))+1,
    year = mdata$Year+1,
    ind = mdata$ind,
    sp = mdata$sp,
    indsp = arrange(unique(mdata[c("ind", "sp")]), ind)$sp,
    dbh = mdata$DBH,
    Np = nrow(eval),
    dbhp = eval$DBH,
    indp = eval$ind,
    yearp = eval$Year+1),
  chains = 2, 
  parallel_chains = 4,
  refresh = 200,
  save_warmup = F
)
fit_sprand <- rstan::read_stan_csv(fit$output_files())
save(fit_sprand, file = "save/species/sprand.Rdata") 
```

```{r sprandmsampling}
load("save/species/sprand.Rdata")  
```

```{r sprandrhat}
mcmc_rhat(rhat(fit_sprand)) 
```

```{r sprandtrace}
mcmc_trace(as.array(fit_sprand, pars = c("gmax[1]", "sigma", "dopt", "ks", "sigmaD", "sigmaK")))
```

```{r sprandpairs}
mcmc_pairs(as.array(fit_sprand, pars = c("gmax[1]", "sigma", "dopt", "ks", "sigmaD", "sigmaK"))) 
```

```{r sprandalpha}
mcmc_intervals(as.array(fit_sprand, pars = "gmax"))
```

```{r sprandpreds}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
dbhtoday <- function(t, gmax, dopt, ks){
  dbh <- 10
  for(i in 1:t)
    dbh <- dbh + gmax*exp(-0.5* (log(dbh / (100*dopt)) / ks)^2)
  return(dbh)  
}
parsi <- as.data.frame(fit_sprand, pars = c("gmax")) %>% 
  reshape2::melt(variable.name = "parameter") %>% 
  group_by(parameter) %>% 
  summarise(value = median(value)) %>% 
  separate(parameter, c("parameter", "ind"), convert = T) %>% 
  reshape2::dcast(ind ~ parameter)
parss <- as.data.frame(fit_sprand, pars = c("dopt_s", "ks_s")) %>% 
  reshape2::melt(variable.name = "parameter") %>% 
  group_by(parameter) %>% 
  summarise(value = median(value)) %>% 
  separate(parameter, c("parameter", "X1",  "sp"), convert = T) %>% 
  reshape2::dcast(sp ~ parameter)
preds <- select(mdata, ind, sp) %>% 
  unique() %>% 
  left_join(parsi) %>% 
  left_join(parss) %>% 
  mutate(Year = list(0:max(eval$Year))) %>% 
  unnest(Year) %>% 
  rowwise() %>% 
  mutate(DBH = dbhtoday(Year+1, gmax, dopt, ks))
ggplot(left_join(mdata, pars), aes(Year, DBH, col = log(gmax), group = ind)) +
  geom_point() +
  geom_point(data = eval, col = "red", size = 2) +
  geom_line(data = preds) +
  viridis::scale_color_viridis() 
```

## Dmax

We used the selected model, *i.e.* Gompertz model [@Herault2011], with $gmax$ as individual fixed effects and $Dopt$ and $Ks$ as individual random effects centered on species fixed effects:

$$  DBH_{t,i,s} \sim \mathcal N  (10 + gmax_i \times \sum _{y=1|DBH_{t=0}} ^{y=t} exp(-\frac12.[\frac{log(\frac{DBH_{t,i}}{100.Dopt_i})}{Ks_i}]^2)), \sigma) \\| Dopt_i \sim \mathcal N(Dopt_s,\sigma_D), Ks_i \sim \mathcal N(Ks_s,\sigma_K) $$

But we constrained $dopt$ fitting $d=\frac{dopt}{DBH_{95}}$ with $d\in[0,1]$.
The model correctly converged ($\hat R < 1.05$) with an acceptable but marked correlation between $dopt$ and $ks$.
$gmax$ posteriors have logical uncertainty but are varying widely among individuals.
Finally predictions of the diameter trajectories by the model are good and realistic.

```{r dmaxfit, eval=F}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
dmax <- cmdstan_model("model/dmax.stan")
fit <- dmax$sample(
  data = list(
    N = nrow(mdata),
    I = max(mdata$ind),
    S = max(mdata$sp),
    Y = max(max(mdata$Year), max(eval$Year))+1,
    year = mdata$Year+1,
    ind = mdata$ind,
    sp = mdata$sp,
    indsp = arrange(unique(mdata[c("ind", "sp")]), ind)$sp,
    dbh = mdata$DBH,
    dmax = summarise(group_by(mdata, sp), dmax = max(DBH))$dmax,
    Np = nrow(eval),
    dbhp = eval$DBH,
    indp = eval$ind,
    yearp = eval$Year+1),
  chains = 2,
  parallel_chains = 4,
  refresh = 200,
  save_warmup = F
)
fit_dmax <- rstan::read_stan_csv(fit$output_files())
save(fit_dmax, file = "save/species/dmax.Rdata")
```

```{r dmaxsampling}
load("save/species/dmax.Rdata")  
```

```{r dmaxrhat}
mcmc_rhat(rhat(fit_dmax))
```

```{r dmaxtrace}
mcmc_trace(as.array(fit_dmax, pars = c("gmax[1]", "sigma", "dopt_s[1]", "ks_s[1]", "sigmaD", "sigmaK")))
```

```{r dmaxpairs}
mcmc_pairs(as.array(fit_dmax, pars = c("gmax[1]", "sigma", "dopt_s[1]", "ks_s[1]", "sigmaD", "sigmaK")))
```

```{r dmaxgmax}
mcmc_intervals(as.array(fit_dmax, pars = "gmax"))
```

```{r dmaxpreds}
mdata <- vroom::vroom("save/species/mdata_species.tsv")
eval <- vroom::vroom("save/species/eval_species.tsv")
dbhtoday <- function(t, gmax, dopt, ks){
  dbh <- 10
  for(i in 1:t)
    dbh <- dbh + gmax*exp(-0.5* (log(dbh / dopt) / ks)^2)
  return(dbh)  
}
pars <- as.data.frame(fit_dmax, pars = c('gmax', "dopt", "ks")) %>% 
  reshape2::melt(variable.name = "parameter") %>% 
  group_by(parameter) %>% 
  summarise(value = median(value)) %>% 
  separate(parameter, c("parameter", "ind"), convert = T) %>% 
  reshape2::dcast(ind ~ parameter)
preds <- pars %>% 
  mutate(Year = list(0:max(eval$Year))) %>% 
  unnest(Year) %>% 
  rowwise() %>% 
  mutate(DBH = dbhtoday(Year+1, gmax, dopt, ks))
ggplot(left_join(mdata, pars), aes(Year, DBH, col = log(gmax), group = ind)) +
  geom_point() +
  geom_point(data = eval, col = "red", size = 2) +
  geom_line(data = preds) +
  viridis::scale_color_viridis() 
```

## Comparisons

The individual random model has the best prediction (lowest RMSEP) associated to the best evaluation (second lowest loo epld), and a decent computing time. 
Moreover, model using species level $Dopt$ and $Ks$ are underestimating $Gmax$ absolute value while overestimating individual variation in $Gmax$ (last figures).
Moreover, constraining $dopt$ fitting $d=\frac{dopt}{DBH_{95}}$ with $d\in[0,1]$ resolved maximum treedepth issues and have similar performances.
This is the best model that I'll use in next steps.

```{r fitssp}
fits <- list(
  "Individual fixed" = fit_indfix,
  "Individual random" = fit_indrand,
  "Species fixed" = fit_spfix,
  "Species random" = fit_sprand,
  "Dmax" = fit_dmax
)
```

```{r spchoicetab}
lapply(fits, mcmc_intervals_data, "RMSEP") %>% 
  bind_rows(.id = "Model") %>% 
  dplyr::rename(RMSEP = m) %>% 
  dplyr::select(Model, RMSEP) %>% 
  left_join(lapply(fits, loo) %>% 
              loo_compare() %>% 
              as.data.frame() %>% 
              rownames_to_column("Model") %>% 
              dplyr::select(Model, elpd_loo)) %>% 
  left_join(lapply(fits, get_elapsed_time) %>% 
              lapply(as.data.frame) %>% 
              bind_rows(.id = "Model") %>% 
              rowwise() %>% 
              mutate(Time = sum(warmup + sample)) %>% 
              dplyr::select(Model, Time) %>% 
              group_by(Model) %>% 
              summarise(Time = max(Time))) %>% 
  kable(caption = "Model choice summary with quality of prediciton (Root Mean Square Error of Prediction, RMSEP), cross-validation (Leave-One-Out Estimate of the expected Log pointwise Predictive Density, LOO ELDP), and speed (Elapsed yime).",
        col.names = c("Model", "RMSEP", "LOO ELDP", "Elapsed time")) 
```

```{r rmsepsp}
lapply(fits, mcmc_intervals_data, "RMSEP") %>% 
  bind_rows(.id = "Model") %>% 
  ggplot(aes(x = Model, xend = Model)) +
  geom_point(aes(y = m), shape = 21, size = 3, alpha = 0.5) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() + ylab("RMSEP") + xlab("") + scale_y_log10() 
```

```{r loosp}
lapply(fits, loo) %>% 
  loo_compare() %>%
  kable() 
```

```{r timesp}
lapply(fits, get_elapsed_time) %>% 
  lapply(as.data.frame) %>% 
  bind_rows(.id = "Model") %>% 
  rowwise() %>% 
  mutate(Time = sum(warmup + sample)) %>% 
  group_by(Model) %>% 
  summarise(Time = max(Time)) %>% 
  ggplot(aes(Model, Time)) +
  geom_point() +
  coord_flip() +
  scale_y_log10() +
  ylab("Time (seconds)") 
```

```{r gmaxsp}
lapply(fits, mcmc_intervals_data, paste0('gmax[', 1:5, ']')) %>% 
  bind_rows(.id = "model") %>% 
  separate(parameter, c("parameter", "individual")) %>% 
  ggplot(aes(x = individual, xend = individual, col = model)) +
  geom_jitter(aes(y = m)) +
  geom_segment(aes(y = ll, yend = hh),
               size = 1, show.legend = F, alpha = 0.5) +
  geom_segment(aes(y = l, yend = h), size = 2, alpha = 0.5) +
  coord_flip() + ylab(expression(G[max])) 
```

```{r lmmsp}
lapply(fits, as.data.frame, pars = 'gmax') %>% 
  bind_rows(.id = "model") %>% 
  reshape2::melt("model", variable.name = "parameter") %>% 
  group_by(model, parameter) %>% 
  summarise(value = median(value)) %>% 
  separate(parameter, c("parameter", "ind"), convert = T) %>% 
  reshape2::dcast(model + ind ~ parameter) %>% 
  left_join(dplyr::select(vroom::vroom("save/species/mdata_species.tsv"),
                          ind, idTree, Family, Genus, species)) %>% 
  unique() %>% 
  group_by(model) %>% 
  do(var = nlme::lme(log(gmax) ~ 1, random=~1|Family/Genus/species, data = .) %>% 
       ape::varcomp(scale = F, cum = F) %>% 
       as.vector() %>% 
       data.frame(level = c("Family", "Genus", "Species", "Individual"), variance = as.vector(.))) %>% 
  unnest(var) %>% 
  select(-`.`) %>% 
  group_by(model) %>% 
  mutate(pct = variance / sum(variance)*100) %>% 
  mutate(level = factor(level, levels = c("Family", "Genus", "Species", "Individual"))) %>% 
  mutate(pct_text = paste0(round(pct), "%")) %>% 
  mutate(pct_text = gsub("^0%", "", pct_text)) %>% 
  ggplot(aes(x = model, y = pct, fill = level)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(y = pct, label = pct_text), col = "white", position = position_stack(vjust = .5)) +
  scale_fill_discrete(expression(sigma^2)) +
  xlab("Model") + ylab("Percentage of variance")  
```

