```{r setupgrowtha, include=FALSE, eval=T}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(bayesplot)
library(rstan)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F, eval=T)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
```

# Growth 1

**Using `Stan` with a lognormal likelihood.**

We used the following growth model with a lognormal distribution to estimate individual growth potential:

\begin{equation} 
  DBH_{y=today,s,i}  - DBH_{y=y0,s,i} \sim \\
  \mathcal{logN} (log(\sum _{y=y_0} ^{y=today} \theta_{1,s,i}.exp(-\frac12.[\frac{log(\frac{DBH_{y,s,i}}{100.\theta_{2,s}})}{\theta_{3,s}}]^2)), \sigma_1) \\ 
  \theta_{1,s,i} \sim \mathcal {logN}(log(\theta_{1,s}), \sigma_2) \\ 
  \theta_{2,s} \sim \mathcal {logN}(log(\theta_2),\sigma_3) \\ 
  \theta_{3,s} \sim \mathcal {logN}(log(\theta_3),\sigma_4)
  (\#eq:gmaxth)
\end{equation} 

In `stan`, we fitted the equivalent model with the following priors:

\begin{equation} 
  DBH_{y=today,s,i}  - DBH_{y=y0,s,i} \sim \\
  \mathcal{logN} (log(\sum _{y=y_0} ^{y=today} \hat{\theta_{1,s,i}}.exp(-\frac12.[\frac{log(\frac{DBH_{y,s,i}}{100.\hat{\theta_{2,s}}})}{\hat{\theta_{3,s}}}]^2)), \sigma_1) \\ 
  \hat{\theta_{1,s,i}} = e^{log(\theta_{1,s}) + \sigma_2.\epsilon_{1,i}} \\ 
  \hat{\theta_{2,s}}  = e^{log(\theta_2) + \sigma_3.\epsilon_{2,s}} \\ 
  \hat{\theta_{3,s}}  = e^{log(\theta_3) + \sigma_4.\epsilon_{3,s}} \\  
  \epsilon_{1,i} \sim \mathcal{N}(0,1) \\
  \epsilon_{2,p} \sim \mathcal{N}(0,1) \\
  \epsilon_{3,p} \sim \mathcal{N}(0,1) \\
  ~ \\
  (\theta_{1,p}, \theta_2, \theta_3) \sim \mathcal{logN}^3(log(1),1) \\ 
  (\sigma_1, \sigma_2, \sigma_3, \sigma_4) \sim \mathcal{N}^4_T(0,1) \\ 
  (\#eq:gmaxstan)
\end{equation} 

```{r samplinga, eval=F}
load("save/mdata.Rdata")
model <- stan_model("model/gmax.stan")
fit <- sampling(model, chains = 2, data = mdata, save_warmup = F, control = list(adapt_delta = 0.99, max_treedepth = 12),
                include = F, pars = c("epsilon_s2", "epsilon_s3", "epsilon_i", "DBH"))
save(data, mdata, fit, file = "model/gmax.Rdata"))
```

```{r fita}
load("model/gmax.Rdata")
```

```{r timea}
get_elapsed_time(fit) %>% 
  as.data.frame() %>% 
  rownames_to_column("chain") %>% 
  mutate(total = warmup + sample) %>% 
  mutate_at(c("warmup", "sample", "total"), lubridate::seconds_to_period) %>% 
  kable(caption = "Elapsed time to fit 187 trees over 37 years and 3 species in the plot 14 of Paracou.")
```

```{r traceplota}
mcmc_trace(fit, pars = c("theta2", "theta3", "sigma[1]", "sigma[2]", "sigma[3]", "sigma[4]", "lp__", 
                         "thetai1[1]", "thetas1[1]", "thetas2[1]", "thetas3[1]"), np = nuts_params(fit))
```

```{r pairsa}
mcmc_pairs(fit, pars = c("theta2", "theta3", "sigma[1]", "sigma[2]", "sigma[3]", "sigma[4]"), np = nuts_params(fit))
```

```{r thetaia, fig.height=12, eval=F}
mcmc_intervals(fit, pars = paste0("thetai1[", 1:2568, "]")) +
  scale_x_log10() +
  theme(axis.text.y = element_blank())
```

