```{r analyses2, include=FALSE, eval=T}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(googlesheets4)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F, eval=T)
```

# Other analyses

In this chapter, I quickly  investigated effects of functional traits and weather on individual growth.

## Methods

I used linear model for individual growth potential relation to functional traits at the individual or species level.
For each dataset, I first used a step procedure to select explanatory variables before using linear models with multiple variables.
I then plotted the individual relations between response and explanatory variables.

```{r dataan2}
growth <- vroom::vroom("save/growthhalf.tsv")
phd <- bind_rows(
  vroom::vroom("data/Measures_Symphonia - AllTraits.tsv") %>% 
    mutate(Genus = "Symphonia") %>%
    rename(Species = Morphotype) %>%
    mutate(Species = ifelse(Species == "Indet.",
                            c("globulifera", "sp.1", "sp.1")[fct_recode(Bark, "globulifera" = "G",
                                                                        "sp.1" =  "S")], Species)),
  vroom::vroom("data/Measures_Eschweilera - AllTraits.tsv") %>%
    filter(!(Plot == 14 & SubPlot == 1 & TreeFieldNum == 760))
) %>% 
  dplyr::select(idTree, SLA, LDMC, LT, LA, CC, brBT, brWD, brBD) %>% 
  left_join(dplyr::select(growth, idTree, species, gmax)) %>% 
  filter(!is.na(gmax)) %>% 
  group_by(species, idTree) %>% 
  summarise_all(mean, na.rm =T)
hydro <- vroom::vroom("data/211124_TraitDatabase_species_means_updatedNames.csv") %>% 
  mutate(species = gsub("_", " ", spName.new)) %>% 
  dplyr::rename(LSWC = LSWC_E.per_corr_mean, RWC = RWC_E.per_corr_mean, Ptlp = ptlp_mean, d13C = d13C.per_mean, CN = C.N_mean,
                stomataD = stomata_density.mm2_mean, gmin = gmin.slope_mean, LWC = LWC_mean, LDMC = LDMC.mg.g_mean,
                N = N.per_mean, C = C.per_mean, P12stem = P12_stem.Mpa_mean, P50stem = P50_stem.MPa_mean, P88stem = P88_stem.MPa_mean,
                P12leaf = P12_leaf.MPa_mean, P50leaf = P50_leaf.MPa_mean, P88leaf = P88_leaf.MPa_mean) %>% 
  dplyr::select(species, LSWC, RWC, Ptlp, d13C, CN, stomataD, gmin, LWC, LDMC, N, C, P12stem, P50stem, P88stem, P12leaf, P50leaf, P88leaf) %>% 
  left_join(group_by(growth, species) %>% summarise(gmax  = median(gmax, na.rm = T)))
fg <- readxl::read_xlsx("data/Appendix_S2-6___S8-13.xlsx", "App.S6-ok", skip = 3) %>% 
  mutate(species = gsub("_", " ", Species)) %>% 
  left_join(group_by(growth, species) %>% summarise(gmax  = median(gmax, na.rm = T))) %>% 
  filter(!is.na(gmax)) %>% 
  dplyr::select(species, gmax, Chlorophyll_content, Thickness, Toughness, Leaf_Area, SLA, C, N, `13C`, Ca, P, K, 
                Trunk_bark_thickness, Sapwood_WSG, WSG, Diameter, SRL, Tissue_density, SRTA, branchiness)
cam <- read_tsv("data/Ziegler2019 - All.tsv") %>% 
  dplyr::select(-Code) %>% 
  reshape2::melt("Species", variable.name = "Trait") %>% 
  mutate(value = gsub("− ", "-", value)) %>% 
  separate(value, c("mean", "sd"), "±", convert = T) %>% 
  mutate(mean = as.numeric(mean)) %>% 
  reshape2::dcast(Species ~ Trait, value.var = "mean") %>% 
  separate(Species, c("Genus", "Species", "Author")) %>% 
  select(-Author) %>% 
  mutate(species = paste(Genus, Species)) %>% 
  left_join(group_by(growth, species) %>% summarise(gmax  = median(gmax, na.rm = T)) %>% separate(species, c("Genus", "Species")))
santi <- read_delim("data/Santiago2018_S1.tsv", delim = " ", skip = 1,
           col_names = c("Genus", "Species", "WD", "X1", "X2",
                         "sapwood_saturated_water_content", "X3", "X4",
                         "sapwood_osmotic_potential_at_full_turgor", "X5", "X6",
                         "sapwood_water_potential_at_turgor_loss_point", "X7", "X8",
                         "total_sapwood_bulk_elastic_modulus", "X9", "X10",
                         "total_sapwood_relative_water_content_at_turgor_loss_point", "X11", "X12",
                         "sapwood_capacitance_at_full_turgor", "X13", "X14")) %>% 
  select(Genus, Species, sapwood_saturated_water_content, sapwood_osmotic_potential_at_full_turgor,
         sapwood_water_potential_at_turgor_loss_point, total_sapwood_bulk_elastic_modulus,
         total_sapwood_relative_water_content_at_turgor_loss_point, sapwood_capacitance_at_full_turgor) %>% 
  left_join(read_delim("data/Santiago2018_S2.tsv", delim = " ", skip = 1,
                       col_names = c("Genus", "Species", "WD", "X1", "X2",
                                     "P50", "X3", "X4",
                                     "slope_P50", "X5", "X6")) %>% 
              select(Genus, Species, P50, slope_P50)) %>% 
  mutate(Genus = recode(Genus, "B." = "Bocoa", "D." = "Dicorynia", "E." = "Eperua",
                        "J." = "Jacaranda", "L." = "Licania", "P." = "Pradosia", "S." = "Sextonia",
                        "T." = "Tachigali", "V." = "Vouacapoua")) %>% 
  mutate(Genus = ifelse(Species == "globulifera", "Symphonia", Genus)) %>% 
  mutate(Genus = ifelse(Species == "sagotiana", "Eschweilera", Genus)) %>% 
  mutate(Genus = ifelse(Species == "persistens", "Lecythis", Genus)) %>% 
  left_join(group_by(growth, species) %>% summarise(gmax  = median(gmax, na.rm = T)) %>% separate(species, c("Genus", "Species"))) %>% 
  filter(!is.na(gmax)) 
isa <- read_tsv("data/fec12452-sup-0002-appendixs1.csv") %>% 
  dplyr::select(Binomial, Pi_tlp) %>% 
  group_by(Binomial) %>% 
  summarise(Pi_tlp = mean(Pi_tlp)) %>% 
  separate(Binomial, c("Genus", "Species")) %>% 
  left_join(group_by(growth, species) %>% summarise(gmax  = median(gmax, na.rm = T)) %>% separate(species, c("Genus", "Species"))) %>% 
  filter(!is.na(gmax)) 
isa2 <- read_tsv("data/fec12452-sup-0002-appendixs1.csv") %>% 
  separate(Binomial, c("Genus", "Species")) %>% 
  group_by(Genus) %>% 
  summarise(Pi_tlp = mean(Pi_tlp)) %>% 
  left_join(separate(growth, species, c("Genus", "Species"))  %>% group_by(Genus) %>% summarise(gmax  = median(gmax, na.rm = T))) %>% 
  filter(!is.na(gmax)) 
guillemot <- read_tsv("data/TAB_FINAL_GUILLEMOTetal.csv") %>% 
 left_join(group_by(growth, species) %>% summarise(gmax  = median(gmax, na.rm = T))) %>% 
    filter(!is.na(gmax)) 
```

## Results

### ParacouITV

Interestingly, leaf dry matter content seems to drive individual growth potential within species of *Symphonia* and *Eschweilera* genera
($\beta=-0.91, p=0.016, R^2=0.083$), but the signal is weak.

```{r, eval=F}
phd %>% 
  mutate(logSLA = log(SLA), logLDMC = log(LDMC), logLT = log(LT), logLA = log(LA),
         logCC = log(CC), logbrBT = log(brBT), logbrWD = log(brWD), logbrBD = log(brBD)) %>% 
  reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
  na.omit() %>% 
  group_by(trait) %>% 
  do(lm(gmax ~ value, data = .) %>% moderndive::get_regression_table()) %>% 
  filter(term != "intercept") %>% 
  dplyr::select(-term) %>% 
  filter(p_value < 0.05)
```

```{r phdlm}
lm(log(gmax) ~ CC + log(LA) + log(LDMC), phd) %>% 
  sjPlot::tab_model(show.icc = F) 
```

```{r phdfig}
ggplot(aes(LDMC, gmax, col = species, group = NA), data = phd) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("Leaf dry matter content (LDMC, g/g)") +  ylab("Individual growth potential (Gmax, cm/yr)") + 
  scale_y_log10() + scale_x_log10() 
```

### hydroParacou

Interestingly, turgor loss point and stomatal density appear to influence species growth potential with a positive effect ($\beta=1.13, p=0.004$ & $\beta=0.28, p<0.001$, $R^2=0.0312$).
Species with more stomata lose their turgidity more quickly but grow faster.

```{r, eval=F}
bind_rows(
  hydro %>% 
    reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
    na.omit() %>% 
    group_by(trait) %>% 
    do(lm(log(gmax) ~ value, data = .) %>% moderndive::get_regression_table()) %>% 
    filter(term != "intercept") %>% 
    dplyr::select(-term) %>% 
    filter(p_value < 0.05),
  hydro %>% 
    reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
    na.omit() %>% 
    group_by(trait) %>% 
    mutate(value = log(abs(value)), trait = paste0("log", trait)) %>% 
    do(lm(log(gmax) ~ value, data = .) %>% moderndive::get_regression_table()) %>% 
    filter(term != "intercept") %>% 
    dplyr::select(-term) %>% 
  filter(p_value < 0.05)
)
```

```{r hydrolm}
lm(log(gmax) ~ Ptlp + log(CN) + log(stomataD),
   mutate_at(hydro, c("Ptlp", "CN", "stomataD"), scale, center = F)) %>% 
  sjPlot::tab_model(show.icc = F)
```

```{r hydrofig}
hydro %>% 
  mutate(logStomataD = log(stomataD)) %>% 
  reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
  na.omit() %>% 
  filter(trait %in% c("Ptlp", "logStomataD")) %>% 
  ggplot(aes(value, gmax)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("") +  ylab("Species growth potential (Gmax, cm/yr)") + 
  scale_y_log10() + facet_wrap(~ trait, scales = "free_x")
```

### Vleminckx et al. 2021

```{r Vleminckxlm}
lm(log(gmax) ~ C + N + K + Sapwood_WSG + WSG, 
   mutate_at(fg, c("C", "N", "K", "Sapwood_WSG", "WSG"), scale)) %>% 
  sjPlot::tab_model(show.icc = F)
```

```{r Vleminckxfig}
fg %>% 
  reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
  na.omit() %>% 
  filter(trait %in% c("N", "K", "Sapwood_WSG", "WSG")) %>%
  ggplot(aes(value, gmax)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("") +  ylab("Species growth potential (Gmax, cm/yr)") + 
  scale_y_log10() + facet_wrap(~ trait, scales = "free_x")
```

### Ziegler et al. 2019

```{r, eval=F}
cam %>% 
  select(-Species, -Genus) %>% 
  reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
  na.omit() %>% 
  group_by(trait) %>% 
  do(lm(gmax ~ value, data = .) %>% moderndive::get_regression_table()) %>% 
  filter(term != "intercept") %>% 
  dplyr::select(-term) %>% 
  filter(p_value < 0.05)
```


```{r Zieglerlm}
lm(log(gmax) ~ branch_vulnerability_slope + HSM_PiTLP_Psi88, cam) %>% 
  sjPlot::tab_model(show.icc = F)
```

```{r Zieglerfig}
cam %>% 
    select(-Species, -Genus) %>% 
  reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
  na.omit() %>% 
  # filter(trait %in% c("N", "K", "Sapwood_WSG", "WSG")) %>%
  ggplot(aes(value, gmax)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("") +  ylab("Species growth potential (Gmax, cm/yr)") + 
  scale_y_log10() + facet_wrap(~ trait, scales = "free_x")
```

### Santiago et al. 2018

```{r, eval=F}
santi %>% 
  mutate(species = paste(Species, Genus)) %>% 
  select(-Species, -Genus) %>% 
  reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
  na.omit() %>% 
  group_by(trait) %>% 
  do(lm(gmax ~ value, data = .) %>% moderndive::get_regression_table()) %>% 
  filter(term != "intercept") %>% 
  dplyr::select(-term) %>% 
  filter(p_value < 0.05)
```


```{r Santiagolm}
lm(log(gmax) ~ total_sapwood_relative_water_content_at_turgor_loss_point, santi) %>% 
  sjPlot::tab_model(show.icc = F)
```

```{r Santiagofig}
santi %>% 
  mutate(species = paste(Species, Genus)) %>% 
  select(-Species, -Genus) %>% 
  reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
  na.omit() %>% 
  mutate(trait = gsub("_", " ", trait)) %>% 
  filter(trait %in% c("total sapwood relative water content at turgor loss point")) %>%
  ggplot(aes(value, gmax)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("") +  ylab("Species growth potential (Gmax, cm/yr)") + 
  scale_y_log10() + facet_wrap(~ trait, scales = "free_x")
```

### Maréchaux et al. 2015


```{r Marechauxlm}
lm(log(gmax) ~ Pi_tlp, isa) %>% 
  sjPlot::tab_model(show.icc = F)
```

```{r Marechauxfig}
ggplot(isa, aes(Pi_tlp, gmax)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("") +  ylab("Species growth potential (Gmax, cm/yr)") + 
  scale_y_log10()
```

```{r Marechaux2lm}
lm(log(gmax) ~ Pi_tlp, isa2) %>% 
  sjPlot::tab_model(show.icc = F)
```

```{r Marechaux2fig}
ggplot(isa2, aes(Pi_tlp, gmax)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("") +  ylab("Species growth potential (Gmax, cm/yr)") + 
  scale_y_log10()
```

### Guillemot et al. 2022

```{r, eval=F}
guillemot %>% 
  select(-genus, -family, -leaf_habit) %>% 
  reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
  na.omit() %>% 
  group_by(trait) %>% 
  do(lm(gmax ~ value, data = .) %>% moderndive::get_regression_table()) %>% 
  filter(term != "intercept") %>% 
  dplyr::select(-term) %>% 
  filter(p_value < 0.05)
```


```{r Guillemotlm}
lm(log(gmax) ~ log(leaf_size) + Leaf_N + Leaf_P + Wood_density, guillemot) %>% 
  sjPlot::tab_model(show.icc = F)
```

```{r Guillemotfig}
guillemot %>% 
  mutate(leaf_size = log(leaf_size)) %>% 
  select(-genus, -family, -leaf_habit) %>% 
  reshape2::melt(c("species", "gmax"), variable.name = "trait") %>% 
  na.omit() %>% 
  mutate(trait = gsub("_", " ", trait)) %>% 
  # filter(trait %in% c("total sapwood relative water content at turgor loss point")) %>%
  ggplot(aes(value, gmax)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab("") +  ylab("Species growth potential (Gmax, cm/yr)") + 
  scale_y_log10() + facet_wrap(~ trait, scales = "free_x")
```
